---
title: "Most Predictive Period"
author: "Mike Schmidt"
date: "8/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(caret)
```

## Load some data to work with 
This data was loaded and saved useing the modified `get_stock()` function from the `test/testing_different_predictions_methos.Rmd` file.  

Here I am loading and subsetting the data to the fields that I will need to test periods. 

```{r}
data<-readRDS('data/AAPL.rds')%>%
  select(date, close, obv:ult_osc, max52week:since52weekLow)
```


I want to test different periods and see if sertain periods are more predictive than others.  

mutate(period_price_diff = c(tail(close, -15), rep(NA, 15)))%>%
        mutate(period_window_perc=(period_price_diff-close)/close*100)%>%
        mutate( test_variable = if_else(period_window_perc<5, 0 , 1, NA_real_ ))%>%
        mutate(ticker = ticker)

```{r}
period_test<-function(stock_data, period_to_test){
  stock_data%>%
    mutate(period_price_diff = c(tail(close, -(period_to_test)), rep(NA, period_to_test)))%>%
    mutate(period_window_perc=(period_price_diff-close)/close*100)%>%
    mutate( test_variable = if_else(period_window_perc<10, 0 , 1, NA_real_ ))
}

rf_data<-period_test(data, 30)%>%
  drop_na()%>%
  select(-close, -period_price_diff, -test_variable)%>%
  select(-roc, -min52occured, -max52occured) ##removed stuff
```

```{r}
set.seed(420)

train<-head(rf_data, round(0.95*nrow(rf_data)))%>%
  select(-date)
test<-tail(rf_data, round(0.05*nrow(rf_data)))
test2<-head(test, 100)
```
## Random Forest
```{r}
train_control <- trainControl(	method="cv",
                                       number=5,
                                       classProbs = TRUE,
                                       savePredictions = TRUE, 
                                       sampling = "down")
rf<- train(
          period_window_perc~ ., 
          data = train, 
          method="rf", 
          trCrontrol = train_control,
          ntree=150, 
          na.action = na.pass, 
          tuneGrid = expand.grid(.mtry = 9), 
          importance = TRUE
        )

rf
varImp(rf)

prediction<-test2%>%
  mutate(prediction = predict(rf, test2), diff = abs(prediction-period_window_perc))

prediction%>%
  select(date, period_window_perc, prediction, diff)%>%
  View()

prediction%>%
  ggplot(aes(x = prediction, y = period_window_perc, size = diff))+
    geom_point(alpha = 0.5)

test

cor(predict(rf, test2), test2$period_window_perc)
```

Seems to predict 50 to 100 periods at about a corr of 0.4. Degrades after that. 





## Neural Network Regression
```{r}
rec_reg <- recipe(period_window_perc ~ ., data = train) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) 
trainX <- train[, -ncol(train)]
trainY <- train$period_window_perc
```


