---
title: "object_detection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown file is following the [object detection tutorial on Rstudio](https://blogs.rstudio.com/tensorflow/posts/2018-11-05-naming-locating-objects/).  I am doing this tutorial because I want to familiarize myself with tensorflow and keras and I also am hoping to be able to count objects for a project I am doing at work which will require going through thousands of images. 

Download instructions for files.  I saved them in the folder `data_1` for this tutorial. 

```
# mkdir data && cd data
# curl -OL http://pjreddie.com/media/files/VOCtrainval_06-Nov-2007.tar
# curl -OL https://storage.googleapis.com/coco-dataset/external/PASCAL_VOC.zip
# tar -xf VOCtrainval_06-Nov-2007.tar
# unzip PASCAL_VOC.zip
# mv PASCAL_VOC/*.json .
# rmdir PASCAL_VOC
# tar -xvf VOCtrainval_06-Nov-2007.tar

```
### Load Packages
```{r}
library(keras)
library(rjson)
library(magick)
library(purrr)
library(tibble)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
```

